---
title: "PCA"
output: pdf_document
---


SITE: http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(factoextra)
```




Iremos analisar a base de dados "meninas_unicas" obtidas do arquivo analise_estrutura.Rmd

```{r}
meninas_unicas <- read.delim("make_binary/meninasbin.csv", stringsAsFactors=FALSE, sep=";")
meninos_bin <- read.delim("make_binary/meninosbin.csv", stringsAsFactors=FALSE, sep=";")
```

```{r}
str(meninas_unicas[, 4:13])
```



```{r}
#any(is.na(log.ir))
#sum( sapply(log.ir, which(is.infinite) )
#dim(log.ir)
```

The function princomp() uses the spectral decomposition approach. The functions prcomp() and PCA()[FactoMineR] use the singular value decomposition (SVD).

vizualizazr os eigenvalues

```{r}
res.pca <- prcomp(meninas_unicas[, 4:14], scale = TRUE)
fviz_eig(res.pca,
         main = "",
         xlab = "Dimensões",
         ylab = "Porcentagem de variáveis explicadas" )
```

Graph of individuals. Individuals with a similar profile are grouped together

```{r}
fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = FALSE     # Avoid text overlapping
             )
```

Grafico das variáveis


```{r}
fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE,     # Avoid text overlapping,
             title = "Variáveis - PCA"
             )
```



```{r}
fviz_pca_biplot(res.pca, repel = FALSE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```


PCA result

```{r}
library(factoextra)
# Eigenvalues
eig.val <- get_eigenvalue(res.pca)
eig.val
```


```{r}
# Results for Variables
res.var <- get_pca_var(res.pca)
res.var$coord          # Coordinates
res.var$contrib        # Contributions to the PCs
res.var$cos2           # Quality of representation 
# Results for individuals
res.ind <- get_pca_ind(res.pca)
res.ind$coord          # Coordinates
res.ind$contrib        # Contributions to the PCs
res.ind$cos2           # Quality of representation 
```


#Variáveis Categóricas
Qualitative / categorical variables can be used to color individuals by groups. The grouping variable should be of same length as the number of active individuals (here 23).

```{r}
groups <- as.factor(meninas_unicas$X.Leaving.option.form.)
fviz_pca_ind(res.pca,
             col.ind = groups, # color by groups
             addEllipses = FALSE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Grupos",
             title = "Individuos - PCA",
             repel = FALSE,
             geom = "point"
             )
```

Calculate the coordinates for the levels of grouping variables. The coordinates for a given group is calculated as the mean coordinates of the individuals in the group.

```{r}
library(magrittr) # for pipe %>%
library(dplyr)   # everything else
# 1. Individual coordinates
res.ind <- get_pca_ind(res.pca)
# 2. Coordinate of groups
coord.groups <- res.ind$coord %>%
  as_data_frame() %>%
  select(Dim.1, Dim.2) %>%
  mutate(competition = groups) %>%
  group_by(competition) %>%
  summarise(
    Dim.1 = mean(Dim.1),
    Dim.2 = mean(Dim.2)
    )
coord.groups
```



```{r}
meninas_leave <- meninas_unicas %>%
  filter(X.Leaving.option.form. > 1)
#meninas_unicas_u$Forma.de.saída.da.opção
res_leave.pca <- prcomp(meninas_leave[, 4:14], scale = TRUE)
fviz_eig(res_leave.pca,
         main = "",
         xlab = "Dimensões",
         ylab = "Porcentagem de variáveis explicadas" )

```

```{r}
fviz_pca_ind(res_leave.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = FALSE     # Avoid text overlapping
             )
```







```{r}
fviz_pca_var(res_leave.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE,     # Avoid text overlapping,
             title = "Variáveis - PCA"
             )
```


```{r}
groups_leave <- as.factor(meninas_leave$X.Leaving.option.form.)
fviz_pca_ind(res_leave.pca,
             col.ind = groups_leave, # color by groups
             addEllipses = FALSE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Grupos",
             title = "Individuos - PCA",
             repel = FALSE,
             geom = "point"
             )
```




```{r}
fviz_pca_biplot(res_leave.pca, 
                # Individuals
                axes = c(2,3),
                geom.ind = "point",
                fill.ind = groups_leave, col.ind = "black",
                pointshape = 21, pointsize = 2,
                palette = "jco",
                addEllipses = TRUE,
                # Variables
                alpha.var ="contrib", col.var = "contrib",
                repel = TRUE
                )
```











